Technique for \earlywarning\ detection of compact binary coalescence and its implication for multi-messenger astronomy

### 2011-06-03 method.tex
Edited section 2.6
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@892 ccdd05d5-56c1-461a-bf98-85f6876523b4

\Table{\label{table:flops}Minimum computational cost of the \textsc{fir}, \textsc{fft} convolution, and \lloid\ methods for a bank of 657 templates sampled at 4096 Hz.}
\br
Method & \flops \\
\mr
\textsc{fir} & $2.4\times10^{13}$ \\
\textsc{fft} convolution & $2.6\times10^8$ \\
\lloid & $4.7\times10^8$ \\
\br
\end{tabular}
\end{indented}
\end{table}

@@

For time slices $s$, the \lloid\ method requires $2 \slicessamps \numsvdtmps f^s$ for evaluating the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ for the linear transformation from the $\numsvdtmps$ basis templates to the $\numtmps$ time-sliced templates, and $\numtmps f^s$ \flops to accumulate the resultant partial \SNR\ stream.

The computational cost of the decimation of the detector data and the interpolation of the partial \SNR\ streams is a little bit more subtle.  Decimation is achieved by applying an \fir\ antialiasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.  Naively, the antialiasing filter with $N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow f^{s-1}$ \flops.  However, it is only necessary to evaluate the antialiasing filter for the fraction $f^s / f^{s-1}$ of the samples that will not be deleted.  Consequently, an efficient decimator that requires only $2 N^\shortdownarrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 N^\shortdownarrow f^s$ \flops\ is possible.

The story is similar for the interpolation filters used to change the sample rates of the partial \textsc{snr} streams.  Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a lowpass filter with $N^\shortdownarrow$.  The lowpass filter requires $2 M N^\shortdownarrow f^{s-1}$ \flops.  However, by taking advantage of the fact that by construction a fraction $f^{s-1}/f$ of the samples are zero, it is possible to build an efficient interpolator that requires only $M N^\shortuparrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 M N^\shortuparrow f^s$ \flops.

>>>>

\editorial{This section's two tables are small.  Any way to put them side by side or next to another figure?}

\Table{\label{table:flops}Minimum computational cost of the \textsc{fir} filter bank, \textsc{fft} convolution, and \lloid\ methods for a bank of 657 templates sampled at 4096 Hz.}
\br
Method & \flops \\
\mr
\textsc{fir} & $2.4\times10^{13}$ \\
\textsc{fft} convolution & $2.6\times10^8$ \\
\lloid & $4.7\times10^8$ \\
\br
\end{tabular}
\end{indented}
\end{table}

@@

For time slices $s$, the \lloid\ method requires $2 \slicessamps \numsvdtmps f^s$ \flops\ 
for evaluating the orthogonal filters, $2 \numtmps \numsvdtmps f^s$ \flops\ for the 
linear transformation from the $\numsvdtmps$ basis templates to the $\numtmps$ time-sliced templates, and $\numtmps f^s$ \flops\ to add the resultant partial \SNR\ stream.

The computational cost of the decimation of the detector data is a little bit more subtle.  Decimation is achieved by applying an \fir\ antialiasing filter and then downsampling, or deleting samples in order to reduce the sample rate from $f^{s-1}$ to $f^s$.  Naively, an antialiasing filter with $N^\shortdownarrow$ coefficients should demand $2 N^\shortdownarrow f^{s-1}$ \flops.  However, it is necessary to evaluate the antialiasing filter only for the fraction $f^s / f^{s-1}$ of the samples that will not be deleted.  Consequently, an efficient decimator that requires only $2 N^\shortdownarrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 N^\shortdownarrow f^s$ \flops\ is possible.

The story is similar for the interpolation filters used to change the sample rates of the partial \textsc{snr} streams.  Interpolation of a data stream from a sample rate $f^s$ to $f^{s-1}$ consists of inserting zeros between the samples of the original stream, and then applying a lowpass filter with $N^\shortdownarrow$ coefficients.  The lowpass filter requires $2 M N^\shortdownarrow f^{s-1}$ \flops.  However, by taking advantage of the fact that by construction a fraction $f^{s-1}/f$ of the samples are zero, it is possible to build an efficient interpolator that requires only $M N^\shortuparrow f^{s-1} \cdot \left( f^s / f^{s-1} \right) = 2 M N^\shortuparrow f^s$ \flops.
